### 2023-12-04 manuscript.tex
Incorporated some suggestions from Max Staller

Therefore, in this work we apply a series of phylogenetic models to a set of orthologous IDRs identified in the \textit{Drosophila} genome to dissect the forces driving their evolution. Our analyses span multiple levels, ranging from the sequences that compose these regions to the distributed features that characterize them as a whole. For the latter, though the previous approach relied on simulations to generate the null distribution for a hypothesis of no constraint, we instead leverage a fully statistical phylogenetic comparative framework~\cite{Cornwell2017}. By comparing models of constrained and unconstrained continuous trait evolution, \textit{i.e.} the Brownian motion and Ornstein-Uhlenbeck models, respectively, we can demonstrate evidence of selective constraint on features independent of any assumptions about the underlying process of sequence evolution. However, we also propose hybrid approaches that combine simulations with phylogenetic comparative methods to test increasingly refined models of IDR evolution. We find that IDRs exhibit unique patterns of amino acid substitution and that in some proteins disorder itself is a dynamically evolving property. Furthermore, though IDRs are broadly unconstrained along several axes of feature evolution, we identify specific clusters of IDRs with shared patterns of constraint. As in yeast, these clusters are enriched for proteins with specific functional annotations, which suggests the preservation of distributed features and their associated functions is a widespread mechanism of IDR evolution.

@@

The first two principal components of the root distributions show little overall structure, though there is a slight enrichment of regions along two axes that correlate with acidic and polar features, respectively (Fig.~\ref{sfig:root}). Likewise, the projections of the rates onto the first two components are largely distributed along the first principal component (Fig.~\ref{fig:rate}A). This and the variable amounts of sequence divergence in the regions led us to suspect the first principal component was a measure of the overall rate of sequence evolution. Plotting the first principal component against the sum of the average amino acid and indel rates as measured by substitution models revealed a strong association (Fig.~\ref{fig:rate}B). We then projected the rates along second and third principal components to determine if these higher order components contained any additional structure. The resulting distribution is roughly triangular and contains three major axes of variation, corresponding to rapid changes in the regions' proportions of acidic, glutamine, and glycine residues (Fig.~\ref{fig:rate}C-D). Inspection of regions selected along these axes confirmed the high rates of evolution of these features (Fig.~\ref{fig:rate}E-G). Furthermore, we observe a similar distribution when the rates of the order regions were projected along their second and third principal components, which suggests a lack of constraint along these axes is a general property of rapidly evolving proteins (Fig.~\ref{sfig:order}).

>>>>

Therefore, in this work we apply a series of phylogenetic models to dissect of evolution of a set of orthologous IDRs identified in the \textit{Drosophila} genome. Our analyses span multiple levels, ranging from the sequences that compose these regions to the distributed features that characterize them as a whole. For the latter, though the previous approach relied on simulations to generate the null distribution for a hypothesis of no constraint, we instead leverage a fully statistical phylogenetic comparative framework~\cite{Cornwell2017}. By comparing models of constrained and unconstrained continuous trait evolution, \textit{i.e.} the Brownian motion and Ornstein-Uhlenbeck models, respectively, we can demonstrate evidence of selective constraint on features independent of any assumptions about the underlying process of sequence evolution. However, we also propose hybrid approaches that combine simulations with phylogenetic comparative methods to test increasingly refined models of IDR evolution. We find that IDRs exhibit unique patterns of amino acid substitution and that in some proteins disorder itself is a dynamically evolving property. Furthermore, though IDRs are broadly unconstrained along several axes of feature evolution, we identify specific clusters of IDRs with shared patterns of constraint. As in yeast, these clusters are enriched for proteins with specific functional annotations, which suggests the preservation of distributed features and their associated functions is a widespread mechanism of IDR evolution.

@@

The first two principal components of the root distributions show little overall structure, though there is a slight enrichment of regions along two axes that correlate with acidic and polar features, respectively (Fig.~\ref{sfig:root}). Likewise, the projections of the rates onto the first two components are largely distributed along the first principal component (Fig.~\ref{fig:rate}A). This observation and the variable amounts of sequence divergence in the regions led us to suspect the first principal component was a measure of the overall rate of sequence evolution. Plotting the first principal component against the sum of the average amino acid and indel rates as measured by substitution models revealed a strong association (Fig.~\ref{fig:rate}B). We then projected the rates along second and third principal components to determine if these higher order components contained any additional structure. The resulting distribution is roughly triangular and contains three major axes of variation, corresponding to rapid changes in the regions' proportions of acidic, glutamine, and glycine residues (Fig.~\ref{fig:rate}C-D). Inspection of regions selected along these axes confirmed the high rates of evolution of these features (Fig.~\ref{fig:rate}E-G). Furthermore, we observe a similar distribution when the rates of the order regions were projected along their second and third principal components, which suggests a lack of constraint along these axes is a general property of rapidly evolving proteins (Fig.~\ref{sfig:order}).
